---
title: "Technical Appendix"
subtitle: "Intergenerational Education Mobility in Uzbekistan"
author: "Research Team"
date: today
format:
  html:
    toc: true
    number-sections: true
  docx:
    toc: true
execute:
  echo: false
  warning: false
  message: false
---

## A. Data Inventory

```{r}
read_csv_safe <- function(path) {
  tryCatch(
    utils::read.csv(path),
    error = function(e) data.frame(note = paste("Could not read", basename(path), "-", conditionMessage(e)))
  )
}

fmt_num <- function(x, digits = 3) {
  sprintf(paste0("%.", digits, "f"), as.numeric(x))
}

fmt_pct <- function(x, digits = 1) {
  sprintf(paste0("%.", digits, "f percent"), 100 * as.numeric(x))
}

fmt_p <- function(x) {
  x <- as.numeric(x)
  if (is.na(x)) {
    return(NA_character_)
  }
  if (x < 0.001) {
    return("<0.001")
  }
  sprintf("%.3f", x)
}

inv_file <- file.path("..", "data", "metadata", "data_inventory_template.csv")
read_csv_safe(inv_file)
```

## B. Variable Dictionary

```{r}
dict_file <- file.path("..", "data", "metadata", "variable_dictionary_template.csv")
read_csv_safe(dict_file)
```

## C. Identification Strategy

- Module A fixed measure set: rank-rank slope, transition matrix, upward mobility rate, downward mobility rate, persistence probability
- Module B: descriptive or associational models with household and regional controls
- Module C: LiTS IV mechanism models (2022-23 only), interpreted as suggestive non-causal evidence
- Modules A and B do not support causal claims

## C1. Module A Locked Measure Set

```{r}
spec_file <- file.path("..", "data", "metadata", "mobility_measure_set.csv")
read_csv_safe(spec_file)
```

## D. Robustness Checklist

- Alternative mobility definitions
- Alternative age windows
- Weighted versus unweighted estimates
- Parent proxy definition checks
- Module C split-threshold sensitivity (`median`, `<=11`, `<=9`)
- Module C sample-composition and fixed-effect attrition checks

## D1. Empirical Audit Summary

```{r}
claim_audit_file <- file.path("..", "outputs", "tables", "empirical_claim_audit.csv")

if (file.exists(claim_audit_file)) {
  claim_audit <- read_csv_safe(claim_audit_file)
  claim_audit$Status <- gsub("_", " ", claim_audit$status)
  knitr::kable(
    claim_audit[, c("module", "check_id", "Status", "detail")],
    col.names = c("Module", "Check", "Status", "Detail")
  )
} else {
  data.frame(note = "Empirical claim-audit outputs are not available yet.")
}
```

## D2. Parent-Measure Robustness

```{r}
parent_robust_file <- file.path("..", "outputs", "tables", "empirical_parent_measure_robustness.csv")

if (file.exists(parent_robust_file)) {
  parent_robust <- read_csv_safe(parent_robust_file)
  parent_robust <- parent_robust[
    parent_robust$subgroup_type == "overall" &
      parent_robust$subgroup_value == "all" &
      parent_robust$metric %in% c("rank_rank_slope", "persistence_probability"),
  ]

  variant_labels <- c(
    baseline_max = "Baseline: max parent category",
    mean_parent_years = "Mean parent years",
    father_only = "Father only",
    mother_only = "Mother only",
    both_parents_observed_only = "Both parents observed only"
  )
  metric_labels <- c(
    rank_rank_slope = "Rank-rank slope",
    persistence_probability = "Persistence probability"
  )

  parent_robust$display <- paste0(
    fmt_num(parent_robust$estimate),
    " [",
    fmt_num(parent_robust$ci_low),
    ", ",
    fmt_num(parent_robust$ci_high),
    "]"
  )
  parent_robust$variant_label <- unname(variant_labels[parent_robust$parent_measure_variant])
  parent_robust$metric_label <- unname(metric_labels[parent_robust$metric])

  parent_robust_wide <- reshape(
    parent_robust[, c("variant_label", "metric_label", "wave_year", "display")],
    idvar = c("variant_label", "metric_label"),
    timevar = "wave_year",
    direction = "wide"
  )
  names(parent_robust_wide) <- sub("^display\\.", "", names(parent_robust_wide))
  parent_robust_wide <- parent_robust_wide[
    order(
      match(parent_robust_wide$variant_label, variant_labels),
      match(parent_robust_wide$metric_label, metric_labels)
    ),
  ]

  knitr::kable(
    parent_robust_wide,
    col.names = c("Parent-measure variant", "Metric", "2010", "2016", "2022-23")
  )
} else {
  data.frame(note = "Parent-measure robustness outputs are not available yet.")
}
```

## D3. Model Inventory

```{r}
model_inventory_file <- file.path("..", "outputs", "tables", "empirical_model_inventory.csv")

if (file.exists(model_inventory_file)) {
  model_inventory <- read_csv_safe(model_inventory_file)
  knitr::kable(
    model_inventory[, c("module", "model", "outcome", "estimator", "family", "link", "weighted", "fixed_effects", "parameter_scale", "interaction_structure")],
    col.names = c("Module", "Model", "Outcome", "Estimator", "Family", "Link", "Weighted", "Fixed effects", "Parameter scale", "Interaction structure")
  )
} else {
  data.frame(note = "Empirical model inventory is not available yet.")
}
```

## E. Diagnostics Outputs

```{r}
mech_sample_file <- file.path("..", "outputs", "tables", "module_c_mechanism_sample.csv")
mech_summary_file <- file.path("..", "outputs", "tables", "module_c_mechanism_summary.csv")
mech_coef_file <- file.path("..", "outputs", "tables", "module_c_mechanism_coefficients.csv")
robust_scenario_file <- file.path("..", "outputs", "tables", "module_c_mechanism_robustness_scenarios.csv")
if (file.exists(mech_sample_file)) {
  read_csv_safe(mech_sample_file)
} else {
  data.frame(note = "Module C mechanism sample diagnostics not available yet.")
}
if (file.exists(mech_summary_file)) {
  read_csv_safe(mech_summary_file)
} else {
  data.frame(note = "Module C mechanism summary outputs not available yet.")
}
if (file.exists(mech_coef_file)) {
  mech_df <- read_csv_safe(mech_coef_file)
  if ("model" %in% names(mech_df)) {
    mech_df[order(mech_df$model), ]
  } else {
    mech_df
  }
} else {
  data.frame(note = "Module C mechanism diagnostics not available yet.")
}
if (file.exists(robust_scenario_file)) {
  read_csv_safe(robust_scenario_file)
} else {
  data.frame(note = "Module C robustness scenarios not available yet.")
}
```

## F. Module C Descriptive Split by Parental Education

```{r}
if (file.exists(mech_summary_file)) {
  group_labels <- c(
    above_median_parent_edu = "Above median parent education",
    lower_or_equal_median_parent_edu = "At or below median parent education"
  )
  outcome_labels <- c(
    education_stopped_covid = "Stopped education due to COVID",
    shared_device = "Child used shared device",
    switched_online = "Switched to online learning",
    challenge_internet = "Internet challenge",
    support_mother = "Learning support from mother"
  )

  module_c_group_table <- read_csv_safe(mech_summary_file)
  module_c_group_table <- module_c_group_table[
    module_c_group_table$group == "parent_education" &
      module_c_group_table$outcome %in% names(outcome_labels),
    c("outcome", "group_value", "estimate", "n_non_missing")
  ]
  module_c_group_table$order_key <- match(
    paste(module_c_group_table$outcome, module_c_group_table$group_value, sep = "::"),
    c(
      "education_stopped_covid::lower_or_equal_median_parent_edu",
      "education_stopped_covid::above_median_parent_edu",
      "shared_device::lower_or_equal_median_parent_edu",
      "shared_device::above_median_parent_edu",
      "switched_online::lower_or_equal_median_parent_edu",
      "switched_online::above_median_parent_edu",
      "challenge_internet::lower_or_equal_median_parent_edu",
      "challenge_internet::above_median_parent_edu",
      "support_mother::lower_or_equal_median_parent_edu",
      "support_mother::above_median_parent_edu"
    )
  )
  module_c_group_table <- module_c_group_table[order(module_c_group_table$order_key), ]
  module_c_group_table$Outcome <- unname(outcome_labels[module_c_group_table$outcome])
  module_c_group_table$Group <- unname(group_labels[module_c_group_table$group_value])
  module_c_group_table$Estimate <- fmt_pct(module_c_group_table$estimate)

  knitr::kable(
    module_c_group_table[, c("Outcome", "Group", "Estimate", "n_non_missing")],
    col.names = c("Outcome", "Group", "Estimate", "Valid N")
  )
} else {
  data.frame(note = "Module C descriptive split outputs are not available yet.")
}
```

## G. Module C Secondary Mechanism Regressions (Appendix Diagnostics Only)

These logit specifications are retained as diagnostics, not as headline evidence. They are reported to document sensitivity to split support and threshold choices in the child-module sample.

```{r}
if (file.exists(mech_coef_file)) {
  mech_term_labels <- c(
    parent_low_edu = "Low parent education",
    urban = "Urban",
    "parent_low_edu:urban" = "Low parent education x urban"
  )
  mech_model_labels <- c(
    m1_switched_online = "Switched to online learning",
    m2_school_closed_no_online = "School closed without online learning",
    m3_education_stopped_covid = "Stopped education due to COVID",
    m4_any_remote_challenge = "Any remote-learning challenge"
  )

  module_c_coef_table <- read_csv_safe(mech_coef_file)
  module_c_coef_table <- module_c_coef_table[
    module_c_coef_table$model %in% names(mech_model_labels),
    c("model", "term", "estimate", "std.error", "p.value")
  ]
  module_c_coef_table$Model <- unname(mech_model_labels[module_c_coef_table$model])
  module_c_coef_table$Term <- unname(mech_term_labels[module_c_coef_table$term])
  module_c_coef_table$Estimate <- fmt_num(module_c_coef_table$estimate)
  module_c_coef_table$`Std. Error` <- fmt_num(module_c_coef_table$std.error)
  module_c_coef_table$`P-value` <- vapply(module_c_coef_table$p.value, fmt_p, character(1))

  knitr::kable(
    module_c_coef_table[, c("Model", "Term", "Estimate", "Std. Error", "P-value")]
  )
} else {
  data.frame(note = "Module C mechanism coefficient outputs are not available yet.")
}
```

## G1. Module C Stoppage Fragility Check

The baseline Module C split uses mean parent years. This is an explicit exception to the main max-parent proxy used in Modules A and B, because the weighted median split becomes support-degenerate under the max-parent rule in the child-module sample. The strict `<=9` split is likewise retained for transparency but treated as a weak-support scenario when the low-parent group becomes too small for stable FE logit estimation.

```{r}
robust_coef_file <- file.path("..", "outputs", "tables", "module_c_mechanism_robustness_coefficients.csv")

if (file.exists(robust_coef_file)) {
  stop_gap_table <- read_csv_safe(robust_coef_file)
  stop_gap_table <- stop_gap_table[
    stop_gap_table$model == "m3_education_stopped_covid" &
      (stop_gap_table$term == "parent_low_edu" | stop_gap_table$status != "estimated"),
    c("scenario_id", "split_rule", "split_threshold", "use_weights", "estimate", "std.error", "p.value", "n_used", "status", "support_reason")
  ]
  scenario_labels <- c(
    baseline_weighted_median = "Weighted median split",
    unweighted_median = "Unweighted median split",
    weighted_leq11 = "Weighted <=11 years split",
    weighted_leq9 = "Weighted <=9 years split",
    weighted_median_maxparent = "Weighted median split (max-parent proxy)"
  )
  stop_gap_table$Scenario <- unname(scenario_labels[stop_gap_table$scenario_id])
  stop_gap_table$Estimate <- fmt_num(stop_gap_table$estimate)
  stop_gap_table$`Std. Error` <- fmt_num(stop_gap_table$std.error)
  stop_gap_table$`P-value` <- vapply(stop_gap_table$p.value, fmt_p, character(1))

  knitr::kable(
    stop_gap_table[, c("Scenario", "split_threshold", "Estimate", "Std. Error", "P-value", "n_used", "status", "support_reason")],
    col.names = c("Scenario", "Threshold", "Log-odds estimate", "Std. Error", "P-value", "Used N", "Status", "Support diagnostics")
  )
} else {
  data.frame(note = "Module C stoppage fragility diagnostics are not available yet.")
}
```

## H. HBS Linkage Diagnostics

The HBS linkage rule is intentionally conservative. An adult aged 25-64 is treated as linkable only if the respondent is coded as a co-resident child of the household head and at least one parental-generation education measure can be recovered from the head/spouse structure with a minimum 12-year age gap. This produces a selective co-resident sample rather than a population-representative intergenerational sample, so the diagnostics are reported before any supplemental HBS model is considered.

```{r}
hbs_linkage_overview_file <- file.path("..", "outputs", "tables", "hbs_linkage_overview.csv")
hbs_linkage_selection_file <- file.path("..", "outputs", "tables", "hbs_linkage_selection.csv")
hbs_linkage_parent_file <- file.path("..", "outputs", "tables", "hbs_linkage_parent_measure_availability.csv")
hbs_linkage_diag_file <- file.path("..", "data", "processed", "hbs_linkage_diagnostics.csv")

hbs_linkage_overview <- if (file.exists(hbs_linkage_overview_file)) read_csv_safe(hbs_linkage_overview_file) else data.frame(note = "HBS linkage overview not available.")
hbs_linkage_selection <- if (file.exists(hbs_linkage_selection_file)) read_csv_safe(hbs_linkage_selection_file) else data.frame(note = "HBS linkage selection diagnostics not available.")
hbs_linkage_parent <- if (file.exists(hbs_linkage_parent_file)) read_csv_safe(hbs_linkage_parent_file) else data.frame(note = "HBS parent-measure availability diagnostics not available.")
hbs_linkage_diag <- if (file.exists(hbs_linkage_diag_file)) read_csv_safe(hbs_linkage_diag_file) else data.frame(note = "HBS linkage diagnostics not available.")

if ("metric" %in% names(hbs_linkage_overview)) {
  overview_table <- hbs_linkage_overview
  overview_table$value <- ifelse(grepl("Percent|Share", overview_table$metric), fmt_pct(overview_table$value), format(as.integer(round(overview_table$value)), big.mark = ","))
  knitr::kable(overview_table, col.names = c("Diagnostic", "Value"))
} else {
  hbs_linkage_overview
}
```

```{r}
if ("metric" %in% names(hbs_linkage_selection)) {
  selection_table <- hbs_linkage_selection
  selection_table$full_sample <- ifelse(selection_table$metric == "Mean age", fmt_num(selection_table$full_sample), fmt_pct(selection_table$full_sample))
  selection_table$linked_sample <- ifelse(selection_table$metric == "Mean age", fmt_num(selection_table$linked_sample), fmt_pct(selection_table$linked_sample))
  knitr::kable(selection_table, col.names = c("Diagnostic", "Group", "Full adult sample", "Linked sample"))
} else {
  hbs_linkage_selection
}
```

```{r}
if ("parent_measure_category" %in% names(hbs_linkage_parent)) {
  parent_table <- hbs_linkage_parent
  parent_table$share <- fmt_pct(parent_table$share)
  knitr::kable(parent_table, col.names = c("Parent-measure category", "N", "Share of adults 25-64"))
} else {
  hbs_linkage_parent
}
```

```{r}
if ("model_justified" %in% names(hbs_linkage_diag)) {
  linkage_note <- if (isTRUE(as.logical(hbs_linkage_diag$model_justified[1]))) {
    "The diagnostics clear the minimum thresholds for a selective supplemental HBS model."
  } else {
    "The diagnostics do not justify supplemental HBS intergenerational estimation: the linked sample is too selective relative to the full adult sample."
  }
  data.frame(note = linkage_note)
} else {
  data.frame(note = "HBS linkage diagnostics summary is not available.")
}
```

## I. HBS Supplemental Co-resident Models (Conditional)

```{r}
hbs_model_status_file <- file.path("..", "outputs", "tables", "hbs_appendix_model_status.csv")
hbs_model_coef_file <- file.path("..", "outputs", "tables", "hbs_appendix_model_coefficients.csv")

hbs_model_status <- if (file.exists(hbs_model_status_file)) read_csv_safe(hbs_model_status_file) else data.frame(note = "HBS appendix model status not available.")
hbs_model_coef <- if (file.exists(hbs_model_coef_file)) read_csv_safe(hbs_model_coef_file) else data.frame(note = "HBS appendix model coefficients not available.")

if ("model_run" %in% names(hbs_model_status) && !isTRUE(as.logical(hbs_model_status$model_run[1]))) {
  data.frame(note = hbs_model_status$note[1])
} else if ("estimate" %in% names(hbs_model_coef)) {
  model_table <- hbs_model_coef[, c("term", "estimate", "std.error", "p.value")]
  model_table$estimate <- fmt_num(model_table$estimate)
  model_table$std.error <- fmt_num(model_table$std.error)
  model_table$p.value <- vapply(model_table$p.value, fmt_p, character(1))
  knitr::kable(model_table, col.names = c("Term", "Estimate", "Std. Error", "P-value"))
} else {
  data.frame(note = "No supplemental HBS co-resident model is reported.")
}
```
