---
title: "Technical Appendix"
subtitle: "Intergenerational Education Mobility in Uzbekistan"
author: "Research Team"
date: today
format:
  html:
    toc: true
    number-sections: true
  docx:
    toc: true
execute:
  echo: true
  warning: false
  message: false
---

## A. Data Inventory

```{r}
read_csv_safe <- function(path) {
  tryCatch(
    utils::read.csv(path),
    error = function(e) data.frame(note = paste("Could not read", basename(path), "-", conditionMessage(e)))
  )
}

inv_file <- file.path("..", "data", "metadata", "data_inventory_template.csv")
read_csv_safe(inv_file)
```

## B. Variable Dictionary

```{r}
dict_file <- file.path("..", "data", "metadata", "variable_dictionary_template.csv")
read_csv_safe(dict_file)
```

## C. Identification Strategy

- Module A fixed measure set: rank-rank slope, transition matrix, upward mobility rate, downward mobility rate, persistence probability
- Module B: fixed-effects models with household and regional controls
- Module C: event-study DiD with region and year fixed effects

## C1. Module A Locked Measure Set

```{r}
spec_file <- file.path("..", "data", "metadata", "mobility_measure_set.csv")
read_csv_safe(spec_file)
```

## D. Robustness Checklist

- Alternative mobility definitions
- Alternative age windows
- Weighted versus unweighted estimates
- Placebo lead terms in event-study
- Region-specific trend sensitivity

## E. Diagnostics Outputs

```{r}
event_file <- file.path("..", "outputs", "tables", "module_c_event_study_coefficients.csv")
if (file.exists(event_file)) {
  event_df <- read_csv_safe(event_file)
  if ("term" %in% names(event_df)) {
    event_df[order(event_df$term), ]
  } else {
    event_df
  }
} else {
  data.frame(note = "Event-study outputs not available yet.")
}
```
