---
title: "Technical Appendix"
subtitle: "Intergenerational Education Mobility in Uzbekistan"
author: "Research Team"
date: today
format:
  html:
    toc: true
    number-sections: true
  docx:
    toc: true
execute:
  echo: true
  warning: false
  message: false
---

## A. Data Inventory

```{r}
read_csv_safe <- function(path) {
  tryCatch(
    utils::read.csv(path),
    error = function(e) data.frame(note = paste("Could not read", basename(path), "-", conditionMessage(e)))
  )
}

inv_file <- file.path("..", "data", "metadata", "data_inventory_template.csv")
read_csv_safe(inv_file)
```

## B. Variable Dictionary

```{r}
dict_file <- file.path("..", "data", "metadata", "variable_dictionary_template.csv")
read_csv_safe(dict_file)
```

## C. Identification Strategy

- Module A fixed measure set: rank-rank slope, transition matrix, upward mobility rate, downward mobility rate, persistence probability
- Module B: descriptive or associational models with household and regional controls
- Module C: LiTS IV mechanism models (2022-23 only), interpreted as suggestive non-causal evidence
- Modules A and B do not support causal claims
- Reform/COVID DiD: optional and conditional on validated admin region-time variation

## C1. Module A Locked Measure Set

```{r}
spec_file <- file.path("..", "data", "metadata", "mobility_measure_set.csv")
read_csv_safe(spec_file)
```

## D. Robustness Checklist

- Alternative mobility definitions
- Alternative age windows
- Weighted versus unweighted estimates
- Parent proxy definition checks
- Module C split-threshold sensitivity (`median`, `<=11`, `<=9`)
- Module C sample-composition and fixed-effect attrition checks

## E. Diagnostics Outputs

```{r}
mech_sample_file <- file.path("..", "outputs", "tables", "module_c_mechanism_sample.csv")
mech_summary_file <- file.path("..", "outputs", "tables", "module_c_mechanism_summary.csv")
mech_coef_file <- file.path("..", "outputs", "tables", "module_c_mechanism_coefficients.csv")
robust_scenario_file <- file.path("..", "outputs", "tables", "module_c_mechanism_robustness_scenarios.csv")
if (file.exists(mech_sample_file)) {
  read_csv_safe(mech_sample_file)
} else {
  data.frame(note = "Module C mechanism sample diagnostics not available yet.")
}
if (file.exists(mech_summary_file)) {
  read_csv_safe(mech_summary_file)
} else {
  data.frame(note = "Module C mechanism summary outputs not available yet.")
}
if (file.exists(mech_coef_file)) {
  mech_df <- read_csv_safe(mech_coef_file)
  if ("model" %in% names(mech_df)) {
    mech_df[order(mech_df$model), ]
  } else {
    mech_df
  }
} else {
  data.frame(note = "Module C mechanism diagnostics not available yet.")
}
if (file.exists(robust_scenario_file)) {
  read_csv_safe(robust_scenario_file)
} else {
  data.frame(note = "Module C robustness scenarios not available yet.")
}
```
