---
title: "Technical Appendix"
subtitle: "Intergenerational Education Mobility in Uzbekistan"
author: "Research Team"
date: today
format:
  html:
    toc: true
    number-sections: true
  docx:
    toc: true
execute:
  echo: false
  warning: false
  message: false
---

## A. Data Inventory

```{r}
read_csv_safe <- function(path) {
  tryCatch(
    utils::read.csv(path),
    error = function(e) data.frame(note = paste("Could not read", basename(path), "-", conditionMessage(e)))
  )
}

fmt_num <- function(x, digits = 3) {
  sprintf(paste0("%.", digits, "f"), as.numeric(x))
}

fmt_pct <- function(x, digits = 1) {
  sprintf(paste0("%.", digits, "f percent"), 100 * as.numeric(x))
}

fmt_p <- function(x) {
  x <- as.numeric(x)
  if (is.na(x)) {
    return(NA_character_)
  }
  if (x < 0.001) {
    return("<0.001")
  }
  sprintf("%.3f", x)
}

inv_file <- file.path("..", "data", "metadata", "data_inventory_template.csv")
read_csv_safe(inv_file)
```

## B. Variable Dictionary

```{r}
dict_file <- file.path("..", "data", "metadata", "variable_dictionary_template.csv")
read_csv_safe(dict_file)
```

## C. Identification Strategy

- Module A fixed measure set: rank-rank slope, transition matrix, upward mobility rate, downward mobility rate, persistence probability
- Module B: descriptive or associational models with household and regional controls
- Module C: LiTS IV mechanism models (2022-23 only), interpreted as suggestive non-causal evidence
- Modules A and B do not support causal claims

## C1. Module A Locked Measure Set

```{r}
spec_file <- file.path("..", "data", "metadata", "mobility_measure_set.csv")
read_csv_safe(spec_file)
```

## D. Robustness Checklist

- Alternative mobility definitions
- Alternative age windows
- Weighted versus unweighted estimates
- Parent proxy definition checks
- Module C split-threshold sensitivity (`median`, `<=11`, `<=9`)
- Module C sample-composition and fixed-effect attrition checks

## E. Diagnostics Outputs

```{r}
mech_sample_file <- file.path("..", "outputs", "tables", "module_c_mechanism_sample.csv")
mech_summary_file <- file.path("..", "outputs", "tables", "module_c_mechanism_summary.csv")
mech_coef_file <- file.path("..", "outputs", "tables", "module_c_mechanism_coefficients.csv")
robust_scenario_file <- file.path("..", "outputs", "tables", "module_c_mechanism_robustness_scenarios.csv")
if (file.exists(mech_sample_file)) {
  read_csv_safe(mech_sample_file)
} else {
  data.frame(note = "Module C mechanism sample diagnostics not available yet.")
}
if (file.exists(mech_summary_file)) {
  read_csv_safe(mech_summary_file)
} else {
  data.frame(note = "Module C mechanism summary outputs not available yet.")
}
if (file.exists(mech_coef_file)) {
  mech_df <- read_csv_safe(mech_coef_file)
  if ("model" %in% names(mech_df)) {
    mech_df[order(mech_df$model), ]
  } else {
    mech_df
  }
} else {
  data.frame(note = "Module C mechanism diagnostics not available yet.")
}
if (file.exists(robust_scenario_file)) {
  read_csv_safe(robust_scenario_file)
} else {
  data.frame(note = "Module C robustness scenarios not available yet.")
}
```

## F. Module C Descriptive Split by Parental Education

```{r}
if (file.exists(mech_summary_file)) {
  group_labels <- c(
    above_median_parent_edu = "Above median parent education",
    lower_or_equal_median_parent_edu = "At or below median parent education"
  )
  outcome_labels <- c(
    education_stopped_covid = "Stopped education due to COVID",
    shared_device = "Child used shared device",
    switched_online = "Switched to online learning",
    challenge_internet = "Internet challenge",
    support_mother = "Learning support from mother"
  )

  module_c_group_table <- read_csv_safe(mech_summary_file)
  module_c_group_table <- module_c_group_table[
    module_c_group_table$group == "parent_education" &
      module_c_group_table$outcome %in% names(outcome_labels),
    c("outcome", "group_value", "estimate", "n_non_missing")
  ]
  module_c_group_table$order_key <- match(
    paste(module_c_group_table$outcome, module_c_group_table$group_value, sep = "::"),
    c(
      "education_stopped_covid::lower_or_equal_median_parent_edu",
      "education_stopped_covid::above_median_parent_edu",
      "shared_device::lower_or_equal_median_parent_edu",
      "shared_device::above_median_parent_edu",
      "switched_online::lower_or_equal_median_parent_edu",
      "switched_online::above_median_parent_edu",
      "challenge_internet::lower_or_equal_median_parent_edu",
      "challenge_internet::above_median_parent_edu",
      "support_mother::lower_or_equal_median_parent_edu",
      "support_mother::above_median_parent_edu"
    )
  )
  module_c_group_table <- module_c_group_table[order(module_c_group_table$order_key), ]
  module_c_group_table$Outcome <- unname(outcome_labels[module_c_group_table$outcome])
  module_c_group_table$Group <- unname(group_labels[module_c_group_table$group_value])
  module_c_group_table$Estimate <- fmt_pct(module_c_group_table$estimate)

  knitr::kable(
    module_c_group_table[, c("Outcome", "Group", "Estimate", "n_non_missing")],
    col.names = c("Outcome", "Group", "Estimate", "Valid N")
  )
} else {
  data.frame(note = "Module C descriptive split outputs are not available yet.")
}
```

## G. Module C Secondary Mechanism Regressions

```{r}
if (file.exists(mech_coef_file)) {
  mech_term_labels <- c(
    parent_low_edu = "Low parent education",
    urban = "Urban",
    gendermale = "Male child",
    "parent_low_edu:urban" = "Low parent education x urban"
  )
  mech_model_labels <- c(
    m1_switched_online = "Switched to online learning",
    m2_school_closed_no_online = "School closed without online learning",
    m3_education_stopped_covid = "Stopped education due to COVID",
    m4_any_remote_challenge = "Any remote-learning challenge"
  )

  module_c_coef_table <- read_csv_safe(mech_coef_file)
  module_c_coef_table <- module_c_coef_table[
    module_c_coef_table$model %in% names(mech_model_labels),
    c("model", "term", "estimate", "std.error", "p.value")
  ]
  module_c_coef_table$Model <- unname(mech_model_labels[module_c_coef_table$model])
  module_c_coef_table$Term <- unname(mech_term_labels[module_c_coef_table$term])
  module_c_coef_table$Estimate <- fmt_num(module_c_coef_table$estimate)
  module_c_coef_table$`Std. Error` <- fmt_num(module_c_coef_table$std.error)
  module_c_coef_table$`P-value` <- vapply(module_c_coef_table$p.value, fmt_p, character(1))

  knitr::kable(
    module_c_coef_table[, c("Model", "Term", "Estimate", "Std. Error", "P-value")]
  )
} else {
  data.frame(note = "Module C mechanism coefficient outputs are not available yet.")
}
```

## H. HBS Linkage Diagnostics

The HBS linkage rule is intentionally conservative. An adult aged 25-64 is treated as linkable only if the respondent is coded as a co-resident child of the household head and at least one parental-generation education measure can be recovered from the head/spouse structure with a minimum 12-year age gap. This produces a selective co-resident sample rather than a population-representative intergenerational sample, so the diagnostics are reported before any supplemental HBS model is considered.

```{r}
hbs_linkage_overview_file <- file.path("..", "outputs", "tables", "hbs_linkage_overview.csv")
hbs_linkage_selection_file <- file.path("..", "outputs", "tables", "hbs_linkage_selection.csv")
hbs_linkage_parent_file <- file.path("..", "outputs", "tables", "hbs_linkage_parent_measure_availability.csv")
hbs_linkage_diag_file <- file.path("..", "data", "processed", "hbs_linkage_diagnostics.csv")

hbs_linkage_overview <- if (file.exists(hbs_linkage_overview_file)) read_csv_safe(hbs_linkage_overview_file) else data.frame(note = "HBS linkage overview not available.")
hbs_linkage_selection <- if (file.exists(hbs_linkage_selection_file)) read_csv_safe(hbs_linkage_selection_file) else data.frame(note = "HBS linkage selection diagnostics not available.")
hbs_linkage_parent <- if (file.exists(hbs_linkage_parent_file)) read_csv_safe(hbs_linkage_parent_file) else data.frame(note = "HBS parent-measure availability diagnostics not available.")
hbs_linkage_diag <- if (file.exists(hbs_linkage_diag_file)) read_csv_safe(hbs_linkage_diag_file) else data.frame(note = "HBS linkage diagnostics not available.")

if ("metric" %in% names(hbs_linkage_overview)) {
  overview_table <- hbs_linkage_overview
  overview_table$value <- ifelse(grepl("Percent|Share", overview_table$metric), fmt_pct(overview_table$value), format(as.integer(round(overview_table$value)), big.mark = ","))
  knitr::kable(overview_table, col.names = c("Diagnostic", "Value"))
} else {
  hbs_linkage_overview
}
```

```{r}
if ("metric" %in% names(hbs_linkage_selection)) {
  selection_table <- hbs_linkage_selection
  selection_table$full_sample <- ifelse(selection_table$metric == "Mean age", fmt_num(selection_table$full_sample), fmt_pct(selection_table$full_sample))
  selection_table$linked_sample <- ifelse(selection_table$metric == "Mean age", fmt_num(selection_table$linked_sample), fmt_pct(selection_table$linked_sample))
  knitr::kable(selection_table, col.names = c("Diagnostic", "Group", "Full adult sample", "Linked sample"))
} else {
  hbs_linkage_selection
}
```

```{r}
if ("parent_measure_category" %in% names(hbs_linkage_parent)) {
  parent_table <- hbs_linkage_parent
  parent_table$share <- fmt_pct(parent_table$share)
  knitr::kable(parent_table, col.names = c("Parent-measure category", "N", "Share of adults 25-64"))
} else {
  hbs_linkage_parent
}
```

```{r}
if ("model_justified" %in% names(hbs_linkage_diag)) {
  linkage_note <- if (isTRUE(as.logical(hbs_linkage_diag$model_justified[1]))) {
    "The diagnostics clear the minimum thresholds for a selective supplemental HBS model."
  } else {
    "The diagnostics do not justify supplemental HBS intergenerational estimation: the linked sample is too selective relative to the full adult sample."
  }
  data.frame(note = linkage_note)
} else {
  data.frame(note = "HBS linkage diagnostics summary is not available.")
}
```

## I. HBS Supplemental Co-resident Models (Conditional)

```{r}
hbs_model_status_file <- file.path("..", "outputs", "tables", "hbs_appendix_model_status.csv")
hbs_model_coef_file <- file.path("..", "outputs", "tables", "hbs_appendix_model_coefficients.csv")

hbs_model_status <- if (file.exists(hbs_model_status_file)) read_csv_safe(hbs_model_status_file) else data.frame(note = "HBS appendix model status not available.")
hbs_model_coef <- if (file.exists(hbs_model_coef_file)) read_csv_safe(hbs_model_coef_file) else data.frame(note = "HBS appendix model coefficients not available.")

if ("model_run" %in% names(hbs_model_status) && !isTRUE(as.logical(hbs_model_status$model_run[1]))) {
  data.frame(note = hbs_model_status$note[1])
} else if ("estimate" %in% names(hbs_model_coef)) {
  model_table <- hbs_model_coef[, c("term", "estimate", "std.error", "p.value")]
  model_table$estimate <- fmt_num(model_table$estimate)
  model_table$std.error <- fmt_num(model_table$std.error)
  model_table$p.value <- vapply(model_table$p.value, fmt_p, character(1))
  knitr::kable(model_table, col.names = c("Term", "Estimate", "Std. Error", "P-value"))
} else {
  data.frame(note = "No supplemental HBS co-resident model is reported.")
}
```
